Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations

### 2021-12-22 analysis/paper/paper.Rmd
draft abundance and gastric results complete

We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during Fall 2019, Fall 2020, and Spring 2021, respectively. The catch per unit effort, measured as the number of Northern Pikeminnow captured per angler hour, for each was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for Fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for Fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for Spring 2021 (Figure \@ref(fig:cpue-fig)). During Fall 2019, we had a total of `r as.english(sum(N_mods[[3]][[1]]$m))` recaptures which included `r as.english(LP_mod[[3]][[1]]$m)` unique individuals (i.e., `r as.english(sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m)` individual was captured more than once). There were `r as.english(sum(N_mods[[3]][[3]]$m))` recaptures in Fall 2020 which included `r as.english(LP_mod[[3]][[3]]$m)` unique individuals. Summaries of parameters for both the single-census and multiple-census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.

Abundance estimates for Northern Pikeminnow in the Deadwater Slough during Fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for Fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")`. In both cases, the multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied (Figure \@ref(fig:abund-fig)). Table \@ref(tab:abund-tab) provides point estimates and confidence intervals for each estimator for the fall mark-recapture sampling events.

All the estimators yielded estimates with overlapping confidence intervals. Because the sampling design most closely matched a multiple census estimator, we feel the Schnabel estimator is the most appropriate. We examined diagnostic plots for that estimator (number of marks vs. proportion of recaptures) and they suggested that we were meeting the assumptions of the Schnabel estimator. Therefore, further results will be based on the Schnabel estimates of abundance.

Using the mean of the Schnabel estimates of Northern Pikeminnow abundance across two fall sampling events, and the mean of the CPUE from those events as well, we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95%CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring of 2020.

@@

We examined the stomach contents of `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow over the course of the study using gastric lavage methods. Of those, we documented some form of stomach contents in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` of the fish and we were able to confirm fish or fish remnants in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")`. Expressed as a percentage, `r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`% of lavaged Northern Pikeminnow had stomach contents of which `r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`% we identified containing fish or fish remnants. Table \@ref(tab:gastric-tab) provides a further breakdown of gastric lavage results for each sampling event.

For Northern Pikeminnow with some form of stomach contents, the average wet weight of total contents was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r prettyNum(min(stmch_df$stomach_contents_weight), digits = 3)`-`r prettyNum(max(stmch_df$stomach_contents_weight), digits = 3)` g). Overall, fish or fish remnants were `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall weight among all stomach contents examined.

>>>>

We caught `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "n_fish"])`, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "n_fish"])`, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "n_fish"])` Northern Pikeminnow during the fall 2019, fall 2020, and spring 2021 efforts, respectively, for a total of `r prettyNum(sum(total_cpue$n_fish))` fish caught using hook-and-line angling. The CPUE, measured as number of Northern Pikeminnow caught per angler hour, was `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2019", "catch_per_hr"], digits = 3)` for fall 2019, `r prettyNum(total_cpue[total_cpue$event_name == "Fall 2020", "catch_per_hr"], digits = 3)` for fall 2020, and `r prettyNum(total_cpue[total_cpue$event_name == "Spring 2021", "catch_per_hr"], digits = 2)` for spring 2021 (Figure \@ref(fig:cpue-fig)). During the mark-recapture efforts, there were `r as.character(english(sum(N_mods[[3]][[1]]$m)))` recaptures in fall 2019 which included `r as.character(english(LP_mod[[3]][[1]]$m))` unique individuals (i.e., `r as.character(english(sum(N_mods[[3]][[1]]$m) - LP_mod[[3]][[1]]$m))` individual was captured more than once). There were an additional `r as.character(english(sum(N_mods[[3]][[3]]$m)))` recaptures in fall 2020 which included `r as.character(english(LP_mod[[3]][[3]]$m))` unique individuals. Summaries of parameters for both the single census and multiple census mark-recapture estimators are provided in Table \@ref(tab:abund-data-lp) and Table \@ref(tab:abund-data-schn), respectively.

Mark-recapture abundance estimates for Northern Pikeminnow in Deadwater Slough during fall 2019 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2019"]), big.mark = ",")` and for fall 2020 ranged from `r prettyNum(min(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` to `r prettyNum(max(all_est$N[all_est$event_name == "Fall 2020"]), big.mark = ",")` (Table \@ref(tab:abund-tab)). In both cases, the multiple census estimators were consistently larger than the single census estimators, although the size of the 95% confidence intervals were more varied (Figure \@ref(fig:abund-fig)). Table \@ref(tab:abund-tab) provides point estimates and confidence intervals for each estimator for the fall mark-recapture efforts.

All mark-recapture estimators yielded estimates with overlapping confidence intervals. However, because our sampling design most closely matched a multiple census estimator (i.e., we released marked fish each day and recaptures from all previous days were recorded for each day), we fell those estimators most appropriate. Additionally, because the Schnabel estimator includes a small sample size modification, as in our case (i.e., few recaptures), we feel it is most appropriate. The Schnabel estimator has the added benefit of more precise confidence intervals and is more straightforward than the Schumacher-Eschmeyer estimate as it is essentially the weighted average among multiple Chapman-modified Lincoln-Peterson estimates. Therefore, further analyses and results will be based on the Schnabel estimates of abundance.

Using the mean of the Schnabel estimates of Northern Pikeminnow abundance across the two fall sampling events (`r prettyNum(mean(all_est$N[all_est$model == "Schnabel"]), big.mark = ",")`) and multiplying the ratio (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"] / mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`) between the spring CPUE (`r round(total_cpue$catch_per_hr[total_cpue$event_name == "Spring 2021"], 2)`) and mean of the fall CPUEs (`r round(mean(total_cpue$catch_per_hr[total_cpue$event_name %like% "Fall"]), 2)`), we estimated there to be `r prettyNum(round(spring_abund$N_S), big.mark = ",")` (95% CI: `r paste(prettyNum(round(c(spring_abund$Lci_S, spring_abund$Uci_S)), big.mark = ","), collapse = " - ")`) Northern Pikeminnow in Deadwater Slough during spring 2021.

@@

We examined the stomach contents of `r prettyNum(gastric_tbl$lavaged[gastric_tbl$event_name == "Total"], big.mark = ",")` Northern Pikeminnow from Deadwater Slough using gastric lavage methods. Of those, we found contents of some form in `r prettyNum(gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$lavaged[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) stomachs and confirmed fish or fish parts in `r prettyNum(gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"], big.mark = ",")` (`r prettyNum((gastric_tbl$fish_contents[gastric_tbl$event_name == "Total"] / gastric_tbl$stmch_contents[gastric_tbl$event_name == "Total"]) * 100, digits = 3)`%) of those. Table \@ref(tab:gastric-tab) summarizes gastric lavage results for each sampling effort. For those individuals with stomach contents, the average wet weight of total contents was `r prettyNum(mean(stmch_df$stomach_contents_weight), digits = 2)` g (median = `r prettyNum(median(stmch_df$stomach_contents_weight), digits = 3)` g; range = `r prettyNum(min(stmch_df$stomach_contents_weight), digits = 3)`-`r prettyNum(max(stmch_df$stomach_contents_weight), digits = 3)` g). Overall, fish or fish remnants consisted of `r prettyNum(mean(stmch_df$p_fish) * 100, digits = 3)`% of the overall of all stomach contents examined.
